\bf Visualising How Non-linear Dimension Reduction Warps Your Data

### 2024-01-22 paper.qmd
Add picorial representation for remove lond edges

```{r}
#| echo: false
#| fig-cap: "Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset. What is the most effective NLDR technique to represent the original data in a 2D space? What is the effective number of bins in each NLDR technique to create a 2D model?"
#| label: fig-diagnosticpltScurve
#| out-width: 100%

@@

Achieving a smooth representation in 2D space is crucial, and one factor influencing this smoothness is the presence of long edges. These long edges occur when a line connects points that are distant in the triangular mesh, impacting the overall smoothness of the mesh.

@@

@@

```{r}
#| echo: false
#| warning: false
#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset with different benchmark values to remove the long edges. What is the effective benchmark value to remove the long edges?
#| label: fig-diagnosticpltScurvelgrm
##| out-width: 100%

>>>>

```{r}
#| echo: false
#| fig-cap: "Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset. What is the most effective NLDR technique to represent the original data in a 2D space? What is the effective number of bins in each NLDR technique to create a 2D model?"
#| label: fig-diagnosticpltScurve
##| out-width: 100%

@@

Achieving a smooth representation in 2D space is crucial, and one factor influencing this smoothness is the presence of long edges (see @fig-modelScurvermlgimp). These long edges occur when a line connects points that are distant in the triangular mesh, impacting the overall smoothness of the mesh.

@@

UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")

trimesh_umap_s_curve <- ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
  geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
  geom_point(size = 2, colour = "#33a02c") +
  coord_equal()

trimesh_umap_s_curve <- trimesh_umap_s_curve +
  #ggtitle("(a)") +
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme_light() +
  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
  ) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3) 
# theme(axis.text = element_text(size = 5),
#       axis.title = element_text(size = 7))

trimesh_gr_umap_s_curve <- colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve, triangular_object = tr1_object_umap_s_curve, distance_col = distance)

trimesh_gr_umap_s_curve <- trimesh_gr_umap_s_curve +
  geom_point(size = 2, colour = "#33a02c") +
  #ggtitle("(b)") +
  xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  theme_light() +
  #coord_equal() +
  theme(legend.position = "none", plot.title = element_text(size = 5, hjust = 0.5, vjust = -0.5),
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()#change legend key width
  ) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

# ### Define type column
# df <- df_all_umap_s_curve |>
#   dplyr::select(tidyselect::starts_with("x")) |>
#   dplyr::mutate(type = "data") ## original dataset
# 
# df_b <- df_bin_umap_s_curve |>
#   dplyr::filter(hb_id %in% df_bin_centroids_umap_s_curve$hexID) |>
#   dplyr::select(-hb_id) |>
#   dplyr::mutate(type = "model") ## Data with summarized mean
# 
# df_exe <- dplyr::bind_rows(df_b, df)
# 
# distance_df_small_edges <- distance_umap_s_curve %>%
#   dplyr::filter(distance < benchmark_umap_s_curve)
# 
# distance_df_long_edges <- distance_umap_s_curve %>%
#   dplyr::filter(distance >= benchmark_umap_s_curve)
# 
# distance_df_edges <- dplyr::bind_rows(distance_df_small_edges, distance_df_long_edges)
# 
# langevitour::langevitour(df_exe[1:(length(df_exe)-1)], lineFrom = distance_df_edges$from , lineTo = distance_df_edges$to, group = df_exe$type, pointSize = 3, levelColors = c("#6a3d9a", "#33a02c"), lineColors = append(rep("black", length(distance_df_small_edges$from)), rep("red", length(distance_df_long_edges$from))))
#

```{r}
#| warning: false
#| echo: false
#| label: fig-modelScurvermlgimp
#| fig-cap: "What does the 2D model, constructed using UMAP (applied to the S-curve dataset), look like in high dimensions with long edges? (<https://youtu.be/pSL_G_n7-fM>) (a) 2D model without removing the long edges, and (b) 2D model with long edges coloured by red. The default benchmark value is used to identify long edges."
##| out-width: 100%

trimesh_umap_s_curve + trimesh_gr_umap_s_curve +
  #plot_annotation(tag_levels = 'a') +
  plot_layout(ncol = 2) &
  theme(legend.position='none', plot.tag = element_text(size = 8))
```

@@

```{r}
#| echo: false
#| warning: false
#| fig-cap: "Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset with different benchmark values to remove the long edges. What is the effective benchmark value to remove the long edges?"
#| label: fig-diagnosticpltScurvelgrm
##| out-width: 100%

### 2024-01-22 paper.tex
Add picorial representation for remove lond edges

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, enhanced, breakable, frame hidden]}{\end{tcolorbox}}\fi

@@

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltScurve-1.pdf}

@@

Achieving a smooth representation in 2D space is crucial, and one factor
influencing this smoothness is the presence of long edges. These long
edges occur when a line connects points that are distant in the
triangular mesh, impacting the overall smoothness of the mesh.

@@

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, interior hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, enhanced]}{\end{tcolorbox}}\fi

@@

{\centering \includegraphics{paper_files/figure-pdf/fig-diagnosticpltScurve-1.pdf}

@@

Achieving a smooth representation in 2D space is crucial, and one factor
influencing this smoothness is the presence of long edges (see
Figure~\ref{fig-modelScurvermlgimp}). These long edges occur when a line
connects points that are distant in the triangular mesh, impacting the
overall smoothness of the mesh.

@@

{\centering \includegraphics{paper_files/figure-pdf/fig-modelScurvermlgimp-1.pdf}

\caption{\label{fig-modelScurvermlgimp}What does the 2D model,
constructed using UMAP (applied to the S-curve dataset), look like in
high dimensions with long edges? (\url{https://youtu.be/pSL_G_n7-fM})
(a) 2D model without removing the long edges, and (b) 2D model with long
edges coloured by red. The default benchmark value is used to identify
long edges.}
